\chapter{Approximation Algorithms and Inapproximability}
\label{cha:approximation}

In this chapter, we investigate the approximability of the \textit{2-Cycle 1-Chain Multi-Donor Kidney Exchange Problem}. We first present a simple approximation algorithm that leverages known results from matching and set packing. We then analyze its approximation ratio and demonstrate that it achieves a constant strictly better than $0.5$.

Based on the two reductions in \autoref{cha:computational_complexity_analysis}, we can also get some results for the hardness of approximation.


\section{Approximation Algorithm for Multi-Donor Exchange}

A natural baseline heuristic is to select only disjoint 2-cycles and ignore all chains. This corresponds to the \emph{alternative donors} model described in \autoref{cha:problem_formulation}, where each patient has one or more donors, but only 2-cycles are allowed. The problem then reduces to maximum matching in an undirected graph, solvable in polynomial time via Edmonds' Blossom algorithm~\cite{edmonds1965paths}. Since each 2-cycle can potentially support up to two chain transplants, this strategy yields a $0.5$-approximation in the worst case.

To improve upon the baseline, we reinterpret our problem as a special case of the \textit{weighted $k$-Set Packing problem}. Details in \autoref{pro:weighted_set_packing}. %, where each feasible exchange, comprising a 2-cycle and up to two chains, is encoded as a subset of at most four patients. And the weight is the number of patients. 

We now describe a hybrid algorithm that combines multiple such approximations to achieve a provable bound strictly better than $0.5$. 
We mentioned a polynomial time algorithm solving \autoref{prob:k_set_packing}, achieving $\frac{3}{k + 1 + \varepsilon}$ approximation ratio for any $\varepsilon > 0$. Based on this, we introduce the following algorithm.

\begin{lemma}
The following algorithm achieves an approximation ratio of at least $\frac{36}{67} - \varepsilon$ for any $\varepsilon > 0$ for the \textit{2-Cycle 1-Chain Multi-Donor Kidney Exchange Problem}.
\end{lemma}

\begin{proof}
To simplify, we assume that $\varepsilon$ is close to $0$. The algorithm considers four steps:
\begin{enumerate}
    \item Use only 2-cycles and apply the Blossom algorithm for exact maximum matching.
    \item Use only 2-cycles with one additional chain (3-element subsets), and apply a $\frac{3}{4}$-approximation algorithm for unweighted 3-Set Packing.
    \item Use only 2-cycles with two chains (4-element subsets), and apply a $\frac{3}{5}$-approximation algorithm for unweighted 4-Set Packing.
    \item Return the best solution among the three.
\end{enumerate}

Let the optimal solution contain:
\begin{itemize}
    \item $a$ subsets of size 4
    \item $b$ subsets of size 3
    \item $c$ subsets of size 2
\end{itemize}
Then:
\[
\mathrm{OPT} = 4a + 3b + 2c
\]
The algorithm achieves:
\[
\mathrm{ALG} = \max\left(\tfrac{12}{5} a,\ \tfrac{9}{4}(a + b),\ 2(a + b + c)\right)
\]
To lower-bound $\mathrm{ALG}$ in terms of $\mathrm{OPT}$, we introduce weights $t_1, t_2 \ge 0$ with $t_1 + t_2 \le 1$, such that:
\[
\begin{aligned}
    \mathrm{ALG} &\ge \tfrac{12}{5} t_1 a + \tfrac{9}{4} t_2 (a + b) + 2(1 - t_1 - t_2)(a + b + c) \\
    &= \left(2 + \tfrac{2}{5} t_1 + \tfrac{1}{4} t_2\right) a
    + \left(2 - 2t_1 + \tfrac{1}{4} t_2\right) b
    + \left(2 - 2t_1 - 2t_2\right) c
\end{aligned}
\]
We seek the largest constant $k$ such that:
\[
\mathrm{ALG} \ge k \cdot \mathrm{OPT} = 4ka + 3kb + 2kc
\]
The corresponding linear program is:
\[
\begin{aligned}
\text{maximize} \quad & k \\
\text{subject to} \quad
& 2 + \tfrac{2}{5} t_1 + \tfrac{1}{4} t_2 \ge 4k \\
& 2 - 2t_1 + \tfrac{1}{4} t_2 \ge 3k \\
& 2 - 2t_1 - 2t_2 \ge 2k \\
& t_1 \ge 0,\ t_2 \ge 0,\ t_1 + t_2 \le 1
\end{aligned}
\]
Solving yields:
\[
\boxed{k = \frac{36}{67} \approx 0.5373}
\]

Note that the approximation algorithm for unweighted 3-Set Packing does not exactly achieve a $\frac{3}{4}$ approximation, but can get arbitrarily close to it. Consequently, our overall algorithm cannot guarantee a $\frac{36}{67}$ approximation ratio, but it can approach this bound arbitrarily closely.

\end{proof}

\section{Hardness of Approximation}


\begin{lemma}
For any fixed constant $n \ge 2$, the maximization version of the \textit{$n$-Cycle Unbounded-Chain Multi-Donor Kidney Exchange Problem} cannot be polynomial time approximated
within $n^{1-\varepsilon}$ for any $\varepsilon  > 0$ unless $\text{P} = \text{NP}$.
\end{lemma}

\begin{proof}[Proof Sketch]
We reduce from the \textit{Longest Path from Fixed Source} problem as shown in \autoref{cha:computational_complexity_analysis}. The number of matched patients is proportional to the length of a simple path in the original graph. Therefore, any approximate solution to the exchange problem would yield a comparable approximation for the longest path problem, violating its known inapproximability bounds.
\end{proof}

\begin{lemma}
The \textit{2-Cycle 1-Chain Multi-Donor Kidney Exchange Problem} is APX-hard.
\end{lemma}

\begin{proof}[Proof Sketch]
We reduce from the \textit{MAX-(2P,2N)-3SAT} problem, which is known to be APX-hard. The construction is the same as the one used in the reduction in \autoref{cha:computational_complexity_analysis}. 
\end{proof}



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../ClassicThesis"
%%% ispell-dictionary: "british" ***
%%% fill-column: 76 ***
%%% End:
